Lattice Boltzmann simulations of magnetically responsive\\ ellipsoidal particle stabilized emulsions

### 2025-03-20 chapters/paper1.tex
Updated conclusion section in aim 1 and fixed non ascii characters in references

In this work, I studied the effect of applied magnetic fields on the
formation of bijels stabilized by magnetic ellipsoidal particles. We
considered oblate, spherical, and prolate particles with a permanent
magnetic dipole moment suspended in a symmetric binary liquid. I found
that, while the overall formation of a bijel is not disrupted by
magnetic interactions, bijels stabilized by oblate or prolate ellipsoids
exhibit anisotropic domain size and an anisotropic tortuosity. For
oblate particles, the domain size increases in the direction
perpendicular to the magnetic field while the tortuosity increases in
the parallel direction. Conversely, for prolate particles, the domain
size increases in the direction parallel to the magnetic field while the
tortuosity increases in the perpendicular direction. Compared with
bijels formed in the absence of a magnetic field, I observed changes in
domain size up to \(30\%\) for non-spherical particles. This effect can
be explained by the re-orientation of the particle dipole axis with the
magnetic field, which in turn re-aligns the interfaces in the direction
of the longer axis of the particles. I further corroborated this
mechanisms by analyzing the coarsening speed and the nematic order of
the particles and the interfaces. I found that jamming of particles is
delayed in the direction of the longer ellipsoidal axis leading to
enhanced coarsening in this direction. The timescale of the delayed
jamming coincides with a slow-down of the orientational ordering of the
particles whereas the alignment of the interfaces increases. I also
found that the angle between the particle dipole axis and the interface
normal changes at this stage. Additionally, I analyzed the radial
distribution function of the particles and found that prolate particles
tend to align side by side, while the oblate particles exhibit a less
regular arrangement. Based on these findings, I propose that bijel
formation in magnetic fields is governed by two coupled effects: During
the initial stage, the magnetic torque rotates the particles to align
the magnetic dipole moment with the field, and the orientational order
of the particles facilitates further coarsening. As coarsening proceeds,
the capillary interactions cause the interfaces to align with the longer
particle axis. The shrinking interface compresses the particles into
ordered arrangements. When jamming sets in, some particles can be forced
to tilt out of the interface and this effect is more prominent for
oblate particles.

>>>>

In this work, I studied the effect of applied magnetic fields on the
formation of bijels stabilized by magnetic ellipsoidal particles. We
considered oblate, spherical, and prolate particles with a permanent
magnetic dipole moment suspended in a symmetric binary liquid. I found
that, while the overall formation of a bijel is not disrupted by
magnetic interactions, bijels stabilized by oblate or prolate ellipsoids
exhibit anisotropic domain size and an anisotropic tortuosity. For
oblate particles, the domain size increases in the direction
perpendicular to the magnetic field while the tortuosity increases in
the parallel direction. Conversely, for prolate particles, the domain
size increases in the direction parallel to the magnetic field while the
tortuosity increases in the perpendicular direction. Compared with
bijels formed in the absence of a magnetic field, I observed changes in
domain size up to \(30\%\) for non-spherical particles. When probing
the differences in the microstructure, I identify that while the
connectivity of the microstructure does not change, 
the time evolution of the area averaged gaussian curvature shows that
orientational ordering of the particles affects the shape of the
coarsening interface. These shape differences are then locked into place
upon jamming, resulting in a more hyperbolic interface in bijels stabilized
by ellipsoidal particles upon application of magnetic fields. The frequency 
of the number of channels changes as a function of distance from the interface. I
characterize this as peak movements closer and further from the interface upon application of magnetic
field, corroborating the microstructure anisotropy characterized.

This effect can
be explained by the re-orientation of the particle dipole axis with the
magnetic field, which in turn re-aligns the interfaces in the direction
of the longer axis of the particles. I further corroborated this
mechanisms by analyzing the coarsening speed and the nematic order of
the particles and the interfaces. I found that jamming of particles is
delayed in the direction of the longer ellipsoidal axis leading to
enhanced coarsening in this direction. The timescale of the delayed
jamming coincides with a slow-down of the orientational ordering of the
particles whereas the alignment of the interfaces increases. I also
found that the angle between the particle dipole axis and the interface
normal changes at this stage. Orientational ordering at the interface was
observed to reduce for disc like particles and increase for rod like
particles upon application of the magnetic field. I analyzed the radial
distribution function of the particles and found that prolate particles
tend to align side by side, while the oblate particles exhibit a less
regular arrangement. Based on these findings, I propose that bijel
formation in magnetic fields is governed by two coupled effects: During
the initial stage, the magnetic torque rotates the particles to align
the magnetic dipole moment with the field, and the orientational order
of the particles facilitates further coarsening. As coarsening proceeds,
the capillary interactions cause the interfaces to align with the longer
particle axis. The shrinking interface compresses the particles into
ordered arrangements. When jamming sets in, some particles can be forced
to tilt out of the interface and this effect is more prominent for
oblate particles.
